---
title: "Reproducible Research Week 2 Peer Review"
author: "Will Kemp"
date: "22 June 2017"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading & Preprocessing the Data

To begin the analysis, I set my working directory and read the data in using the 'read.csv' function. I then convert the date column to a POSIXCT format to allow easier analysis later on.

```{r read}
setwd("C:/Users/william.kemp/Desktop/Data Science - John Hopkins/5. Reproducible Research/Week 2 Peer Review/Data")
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date,format ="%Y-%m-%d")
```

## What is mean total number of steps taken per day?

Firstly, I will remove the missing values from this dataset. Then I aggregate the number of steps taken per day using.

Using this new aggregated data I can calculate the mean number steps taken per day to be:

```{r aggsteps}
activity2 <-  activity[is.na(activity$steps) == FALSE,]
aggdata <- aggregate(activity2$steps, by = list(activity2$date),FUN = sum)
```

Histogram of the number of steps per day:
```{r histsteps}
hist(aggdata$x,freq=T, main = "Histogram of Steps (per Day)", xlab = "Steps")
```

The mean and median steps per day:
```{r meanmedsteps}
mean(aggdata$x,na.rm =T)
median(aggdata$x, na.rm=T)
```


## What is the average daily activity pattern?

Aggregating the dataset (with NAs removed) by the time interval I plot a timeseries of the average number of steps taken per 5 minute interval

```{r plotint, echo=TRUE}
aggdata <- aggregate(activity2$steps, by=list(activity2$interval),FUN = mean)
plot(aggdata$Group.1,aggdata$x,type = "l",xlab ="Interval",ylab="Average Steps", main = "Average Steps per Interval")
```


The interval containing the max. number of steps is:
```{r maxint, echo=TRUE}
subset(aggdata,max(aggdata$x)==x)$Group.1
```

## Imputing missing values

The total number of missing values from the original dataset are:

```{r totmissing, echo=TRUE}
sum(is.na(activity))
```

To complete these missing values I will use a linear regression model to estimate the missing data:
```{r lmmissing, echo=TRUE}
model <- lm(steps~date + interval, data = activity)
activity <- within(activity,steps <- ifelse(is.na(steps),as.integer(predict(model,activity[,c(2,3)])),steps))
aggdata <- aggregate(activity$steps, by = list(activity$date),FUN = sum)
```

I then aggregate the data by day as before, and create a histogram of the steps taken, with the new completed dataset:

```{r histcompletedat, echo=TRUE}
aggdata <- aggregate(activity$steps, by = list(activity$date),FUN = sum)
hist(aggdata$x,freq=T, main = "Histogram of Steps (per Day)", xlab = "Steps")
```

The mean and median of the new dataset are as follows:

```{r meanmedcompletedat, echo=TRUE}
mean(aggdata$x,na.rm =T)
median(aggdata$x, na.rm=T)
```

Imputing the missing data using a linear model has had the impact of slightly reducing both the mean and median of the data.

## Are there differences in activity patterns between weekdays and weekends?

Firstly, I create a new factor variable using the weekdays() function. I then change these to the labels 'Weekend' and Weekday:

```{r weekday, echo=TRUE}
activity$weekday <- weekdays(activity$date)
activity$weekday[!(activity$weekday %in% c("Saturday","Sunday"))] <- "Weekday"
activity$weekday[activity$weekday %in% c("Saturday","Sunday")] <- "Weekend"
activity$weekday <- as.factor(activity$weekday)
```

I then aggregate the number of steps by the newly created factor variable. Finally, I plot a lattice time series graph to show the difference in evolution of steps per 5 minute interval

```{r weekdayplot, echo=TRUE}
require(lattice)
aggdata <- aggregate(activity$steps, by=list(activity$weekday,activity$interval),FUN = mean)
xyplot(aggdata$x~aggdata$Group.2|as.factor(aggdata$Group.1),type="b",layout =c(1,2))
```

